---
title: "Assignment 4 - Task 2"
author: "Evie Vermeer"
date: "11/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(here)
library(tidyverse)
library(ggplot2)
library(janitor)
library(kableExtra)
library(lubridate) 
library(plotly)
```

```{r, include=FALSE}
hares <- read_csv(here("data","bonanza_hares.csv"))
```

### Introduction 

A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last so you know what’s contained in the report.

### Data and Analyses 
A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.

### Annual Juvenile Hare Trap Counts 

```{r}
juvenile_hares <- hares %>% 
  filter(age=="j") %>% 
  get_dupes(trap) %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year = year(date_new)) %>% 
  group_by(year)

ggplot(data=juvenile_hares, aes(x=year)) +
  geom_histogram(bins=14,color="black", fill="burlywood4")+
  scale_x_continuous(breaks = 1999:2012)+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(title="Annual Juvenile Hare Trap Counts",
       x="Year",
       y="# of Hares")
  
#Need to create x-axis scale with years 

```
### Figure caption and takeaways



Takeaways 




```{r}

ggplot(data=juvenile_hares, aes(x=weight, fill=sex, color=sex))+
  theme_classic()+
  geom_histogram(alpha=.2,color="white")+
  scale_fill_manual(values=c("f"="magenta","m"="navy","NA"="gray"))+
  facet_wrap(~grid)
  

```


