---
title: "Assignment 4 - Task 2"
author: "Evie Vermeer"
date: "11/4/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(here)
library(tidyverse)
library(ggplot2)
library(janitor)
library(kableExtra)
library(lubridate) 
library(plotly)
```

```{r, include=FALSE}
hares <- read_csv(here("data","bonanza_hares.csv"))
```

### Introduction 

A 4 - 5 sentence "mini-introduction" to what this study is exploring, including basic background information about the subject, site and study (with citations as needed), and a brief overview of what is included in the exploratory report. You should probably do this last so you know what’s contained in the report.

### Data and Analyses 
A 3 - 4 sentence section in which you briefly describe the data, including citation & credit as needed, and provide an overview of the main analysis, statistical methods and tools used.

### Annual Juvenile Snowshoe Hare Trap Counts 

```{r}
juvenile_hares <- hares %>% 
  filter(age=="j") %>% 
  get_dupes(trap) %>% 
  mutate(date_new = mdy(date)) %>% 
  mutate(year = year(date_new)) %>% 
  group_by(year) %>% 
  mutate(site = case_when(
    grid %in% c("bonrip") ~ "Bonrip",
    grid %in% c("bonmat") ~ "Bonmat",
    grid %in% c("bonbs") ~ "Bonbs"))

ggplot(data=juvenile_hares, aes(x=year)) +
  geom_histogram(bins=14,color="black", fill="burlywood4")+
  scale_x_continuous(breaks = 1999:2012)+
  theme_classic()+
  theme(plot.title = element_text(hjust = 0.5))+
  labs(title="Annual Juvenile Snowshoe Hare Trap Counts",
       x="Year",
       y="# of Hares")
```
### Figure caption and takeaways



Takeaways 




```{r}

ggplot(data=juvenile_hares, aes(x=weight, fill=sex, color=sex))+
  theme_classic()+
  geom_histogram(alpha=.2,color="white")+
  scale_fill_manual(values=c("f"="magenta","m"="navy","NA"="gray1"),labels=c("Female","Male","NA"))+
  facet_wrap(~site,)+
   labs(x = "Weight (g)",
       y= "# of Hares", 
       fill = "Sex",
       title = "Juveline Snowshoe Hare Weights by Sex and Site")+
  theme(plot.title = element_text(hjust = 0.5), strip.background = element_rect(color="white", fill="gray"))
  
```


```{r}
# Add table caption

juvenile_hare_site_stats <- juvenile_hares %>% 
  mutate(sex_full = case_when(
    sex %in% c("f") ~ "Female",
    sex %in% c("m") ~ "Male")) %>% 
  group_by(site,sex_full) %>% 
  summarize(mean_weight=mean(weight,na.rm=TRUE),
            sd_weight=sd(weight, na.rm=TRUE),
            sample_size=n()) 
 
juvenile_hare_site_stats %>% 
  kable(col.names = c("Site",
                      "Sex",
                      "Average Weight (g)",
                      "Standard Deviation (g)",
                      "Sample Size (n)"),
        caption = "Average Weights of Juvenile Snowshoe Hares by Site and Sex",
        digits = 1) %>% 
  kable_styling(full_width = FALSE) 
                
                  
# How to remove NaN? 
# Is there a way to facet a table?
```

```{r}
### Add table caption 

juvenile_hare_stats <- juvenile_hares %>% 
  mutate(sex_full = case_when(
    sex %in% c("f") ~ "Female",
    sex %in% c("m") ~ "Male")) %>% 
  group_by(sex_full) %>% 
  summarize(mean_weight=mean(weight,na.rm=TRUE),
            sd_weight=sd(weight, na.rm=TRUE),
            sample_size=n()) 
 
juvenile_hare_stats %>% 
  kable(col.names = c("Sex",
                      "Average Weight (g)",
                      "Standard Deviation (g)",
                      "Sample Size (n)"),
        caption = "Average Weights of Juvenile Snowshoe Hares by Sex",
        digits = 1) %>% 
  kable_styling(full_width = FALSE)
                      

```

```{r}

ggplot(data=juvenile_hares, aes(x=weight))+
  geom_histogram(bins=18)+
  facet_wrap(~sex)

ggplot(data=juvenile_hares, aes(sample=weight)) +
  geom_qq()+
  facet_wrap(~sex)

```

```{r}
male_juvenile_hares <- juvenile_hares %>% 
  filter(sex == "m") %>% 
  pull(weight)
  
female_juvenile_hares <- juvenile_hares %>% 
  filter(sex == "f") %>% 
  pull(weight)  
  
sex_t <- t.test(female_juvenile_hares,male_juvenile_hares, na.rm=TRUE)

sex_cohen <- cohen.d(female_juvenile_hares,male_juvenile_hares, na.rm = TRUE)

sex_cohen
sex_t
```





